\documentclass[a4paper,11pt]{article}
%\documentclass[a4paper,draft]{jpconf}

\input csz

\usepackage{fullpage}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}

\def\smalltt#1{{\small\texttt{#1}}}
\def\foottt#1{{\footnotesize\texttt{#1}}}

\def\q{\quad}

\def\tightlist{\setlength{\topsep}{0pt}\setlength{\itemsep}{0pt}}

\def\greed{\textsc{Greed}\xspace}
\def\greedhome{\textsc{Greed}@\texttt{home}\xspace}
\def\root{{ROOT}\xspace}
\def\xrootd{\texttt{xrootd}\xspace}
\def\aliroot{{AliROOT}\xspace}
\def\alien{{AliEn}\xspace}
\def\gled{\textsc{Gled}\xspace}
\def\web{\texttt{web}\xspace}
\def\grid{\texttt{grid}\xspace}
\def\bittorrent{\texttt{BitTorrent}\xspace}


\title{Greed Manifesto}

\author{Matev\v{z} Tadel \& Jan Jona Javor\v{s}ek\\
        {\small written by Matev\v{z} Tadel}}

\date{\today}


\begin{document}

\maketitle

\setcounter{section}{-1}


\section{Document status}

\noindent\textbf{v1.0, Oct--Dec 2007.}
Initial version intended for friends. Presents a dream of the \greed
project and wanders off into the realms of eventual possibilities.

\smallskip

\noindent\textbf{v1.0.1, 20 Jan 2008.}
Incorporated comments from Jan Jona.

\smallskip

\noindent\textbf{v1.1.0, 16 Nov 2008.}
Remove section \emph{"Technological overview"} and \emph{"Work-plan
  towards minimal implementation"}. They are now available from the
web -- as they should really be discussed further.


% ========================================================================

\section{Introduction}

\greed is a vehicle for dissemination and universal exploitation of
scientific and educational programs. It combines the concepts of
\grid, \texttt{computing@home} and massive multi-player online games
with the \emph{free and open-source software} ideology and development
practices. The main aim of \greed is to become a link between
institutions for research \& education and Internet users by providing
a global and open computing environment that allows all parties to
pursue their own goals and interests. Institutions utilize the
computing resources of Internet users and in return provide a
high-reliability server infrastructure needed for operation of
persistent virtual worlds created and maintained by communities of
\greed members.

In its essence, \greed is a hack to make this world a better place.
It attempts to expose greed both as the ultimate source of motivation
as well as an unwholesome tendency that limits the human cooperation
and hinders global development. But greed is not a very good name for
a project, unless it is an acronym: \greed is a \emph{Global Research
  Environment for Equitable Development}.\footnote{\emph{Global
    Research Environment for Eidetic Design} is also an option,
  suggestions are welcome.}


% ========================================================================

\section{Roots of the \greed initiative}

Every global initiative starts with a concrete problem and with a
vision of how the present elements can be combined for the common
good.  \greed is no exception to this rule: it is motivated by the
computing needs of the \emph{Large Hadron Collider} (LHC) experiments
and by the desire of its individual members to expose the computing
technologies developed for operation of LHC and use them as an
outreach stunt for popularization of scientific theories and
contemporary technological endeavors undertaken for their exploration.

While the basic computing requirements of the LHC experiments are
covered by the dedicated infrastructure of the LHC computing \grid,
there is little spare capacity for advanced searches for exotic
particles and new physics which require significant computing
resources, mainly in terms of event simulation and analysis of both
simulated and real experiment data.\footnote{The resources provided to
  the ALICE experiment during the first four years of the detector
  operation are actually significantly below the pledged level.}
The additional computing resources provided by the public at large
could thus contribute the final stone to the full exploitation of the
physics potential of the LHC.

As the full experiment software and accompanying data would have to be
distributed for the operation of such infrastructure, it would also
become feasible for anybody to run simple data visualization and
analysis applications allowing better understanding of the LHC
project, of the LHC experiments and of high-energy physics in general.
This is of particular interest for universities and high-schools as
the programs could be used for student projects or for exercises in
physics, computing and statistics. Additionally, educational
institutions usually own a fair number of computers (desktops in
offices and computing rooms) that are only utilized half of the time
as well as have a reasonable WAN up-link and can thus also offer CPU
power for utilization by the LHC experiments.

On the other side of the equation we have 1.2 billion Internet
users that we want to inform about and engage in the latest scientific
and technological developments in an attractive and accessible manner.
This calls for an analysis of the possible user base, keeping in mind
that another purpose of the \greed is to harvest the computing
resources of its target audience for scientific computation.


\subsection{Possible audiences of the \greed}

The success of several \smalltt{computing@home} projects during the
last five years\footnote{Currently folding@home is the most active
  one.  In Dec 2007 it passed the tera-flop mark with 500k registered
  users. During year 2007 PS3 clients accounted for a better part
  of the project's growth.} points to a surprisingly high level of
interest for scientific computing present among the general public.
The basic terms of contract between the users and the project seem, in
most cases, deceptively simple. The project provides a computing
platform and a basic community portal that often consists mostly of
expert-level information (references to scientific papers or research
laboratory web-pages). Based on this rudimentary infrastructure, the
users connect into the distributed computing environment, contributing
their CPU, disk-space and bandwidth for no apparent benefit. But
people who are willing to go through the trouble of installing the
computing client and paying for the electricity\footnote{Estimating
  power consuption to 100W per core, the monthly consumption is 48kWh.
  For France this makes about 6.5 EUR per month per core.} more than
likely care about what they are contributing to, even though this can
be any of the following things:
\begin{itemize}\tightlist
\item rational exploitation of existing computing resources;
\item development and utilization of cool new software and ideas
  pertaining to networks, computing and information technologies in
  general (e.g. FOSS movement, \grid computing, peer-to-peer
  computing);
\item supporting a specific project due to its perceived
  scientific or humanitarian merit.
\end{itemize}
Whatever are the actual motives, the contributors are effectively
forming a community and many of them would feel inclined towards
investing some of their time into further exploration of the project
they are contributing to. The idea of \greed encompasses all of the
above motivations and it would almost certainly appeal to the majority
of existing contributors.

At this point let us also consider the popularity of community
knowledge sites (Wikipedia), of alternate reality worlds (SecondLife)
and of massive multi-player online games (World of Warcraft, EVE
Online). Most of the services in the last two categories also require
a subscription with monthly charges up to 25\$. There are two
common attributes of users who participate in these activities that
also make them possible \greed contributors:
\begin{itemize}\tightlist
\item they have a PC with Internet access and use it as a communication
  engine, as a hobby, or simply for fun;
\item they spend significant amounts of their time online and are
  willing to contribute their work or effort to the community.
\end{itemize}
The combined user bases of these projects well exceed 20 million
people. But unlike the contributors to \smalltt{computing@home}
projects, the users of these community-engines do not have a common
set of motives. Thus either a sub-set of of users needs to be selected
or, alternatively, a variety of strategies need to be employed to
arouse the interest of as many users as possible.

To summarize, there are many existing activities on the Internet that
provided aspiration for \greed. One could even say that \greed is a
peculiar combination of elements from the projects mentioned in this
section that have not been placed into a common context yet.


\subsection{High-level description of \greed worlds}

Persistent multi-user virtual worlds have been chosen as the primary
means of dissemination because they appeal to a variety of audiences,
from primary school level onwards, determined by the complexity and
scope of the virtual world elements. To maintain coherency and allow
information exchange between different worlds, a common basic world
structure must be employed. It should be general enough to allow a
multitude of scientific and technological elements to be combined into
coherent worlds of different characters. At the same time, within
the mechanics of each such world, it must be possible for users to
participate in the world development on completely different levels of
engagement and personal time investment.

To achieve that, the worlds are presentable on different levels. Each
presentation offers a specific view of the world and offers a
specialized user interface for interaction with the presented world
elements.
\begin{itemize}\tightlist

\item \emph{Strategic} view provides insight into relations among
  entities on a large-scale by presenting status and activity of
  large, compound world entities and their communications. This
  information can be combined with maps or diagrams of extensive
  sections of space, like galactic regions, solar or planetary
  systems, or individual planets.  When presenting a specific element,
  e.g., production, distribution and consumption of a valuable
  resource, the relation among contributing entities can be visualized
  as a 3-dimensional graph, supplemented by histograms, plots and
  tables. High-level settings can be changed here, affecting a
  multitude of entities by effectively altering the parameters of
  world development algorithms.

\item \emph{Operational} view focuses on specific tasks or on
  relations among entities in a more local context. For example, this
  view would be used for planning and execution of a
  resource-discovery mission followed by construction of
  resource-extraction and transport facilities. Settings from the
  operational level influence a specific group of entities via
  interaction with the control agents, not with the individual
  entities. As above, map and 3D-graph views can be employed. In this
  mode it is also possible to use an immersive 3D-view, augmented with
  overlays of schematic data.

\item \emph{Tactical} view presents environment and world entities in
  detail and allows management of individual entities by assigning them
  specific tasks or control agents. Map and augmented 3D views are
  used here, possibly accompanied by a radar-like view of the
  operational area.

\item \emph{Entity} view focuses on different aspects of control over
  an individual entity. This ranges from first-person view and
  operation of a vehicle or a space-ship to control over production
  parameters of a factory or laboratory.

\end{itemize}
Strategic and operational views can focus on different world aspects,
e.g., on transport routes, on energy acquisition and distribution
networks, on world exploration and surveillance, or on progress in
science and engineering.

The initial layout of the virtual worlds follows the structure of a
solar system\footnote{Different topologies with similar consequences
  for the world dynamics are also possible, e.g. \emph{Emental} world
  consisting of a number of Dyson's spheres connected by a set of
  stable tubular wormholes.} and prompts users to engage in
exploration, resource acquisition, infrastructure construction and
research activities by controlling the in-world entities. The initial
time is set about 20 years into the future, with reasonably
extrapolated technologies and with the addition of two new fictitious
technologies: the quark-gluon plasma reactor and the Higgs-field
driven vacuum polarization chamber, both spin-offs of the LHC project,
allowing construction of a new generation of extreme energy sources,
control over large-scale quantum coherence and rudimentary
manipulation of the space-time continuum.

To underline the fragility of our civilization, the story could begin
with a cataclysm: human population of Earth is decimated by a global
disaster (caused by biological agents, comet impact, or global
redefinition of weather patterns) and the survivors are in position to
operate robotic factories and agents. The resources from Earth and
Moon are used to build Foundation vessels that are sent through poorly
controlled wormholes into random solar systems of our universe where
they have to bootstrap new civilizations. Of course, the users
themselves will determine for each individual world if its further
development will be carried out in a cooperative or in an opposing
manner.

Humans are a scarce and vulnerable resource in these conditions,
vitally needed for scientific and technological development as well as
for operation, maintenance and upgrades of existing technological
units. The human population dynamics and living conditions of the
humans determine the number of skilled scientists and engineers that
are produced within each simulated community and become available as
controllable world entities.

To actively employ such a specialist, the user must also spend
\emph{research credits} that can only be gained by providing computing
resources to \greed institutions. Alternatively, some research credit
can also be gained via online quizzes where a user is given a specific
data-set on which he must perform data-analysis and correctly reply to
a set of questions. The virtual-world research cost needed to achieve
major scientific and technological breakthroughs far exceeds the
computing resources of a single user, thus forcing the users to form
cooperative communities.

Introduction of new technologies and research paths is determined both
by the world maintainers and by the users' communities. Users
implicitly provide the development preferences by concentrating their
effort on specific areas while the world maintainers need to provide
adequate scientific and technological improvements. New technologies
will inevitably also introduce new world entities and provide new
requirements for extensions of the world mechanics. At this point the
developers of the core framework and world steering algorithms will
have to incorporate the new technology into the system and make sure
that it functions consistently with the existing world elements.

The in-world technologies introduced in the future will eventually
allow controlled long-range space travel, leading to possibilities for
further exploration and colonization as well as for establishment of
connections between different worlds, to increased cooperation and
possibly to conflict.


\subsection{Possible extensions of \greed world mechanics}

This section presents a review of different research fields that could
be introduced into \greed worlds either to increase their level of
detail or to imbue them with realistic scientific models and knowledge
databases. In this way \greed can provide a virtual environment that
is based on the latest research results and can be directly used not
only as a tool for teaching and for individual exploration of recorded
knowledge but also as an engine for simulation of complex systems. In
this spirit it is hoped that other research institutions and projects
will join the \greed initiative by providing their knowledge, manpower
and server infrastructure to receive the benefits of user-provided CPU
power and popularize their field of expertise among the general
public.

Astronomy and cosmology strive for the understanding of the universe
and of all the phenomena that proceed within it on a wide range of
spacetime-scales.  It is therefore necessary to include the
astronomical facts, at least at a basic level, into the world
mechanics from the very beginning.  With the interest from the
astro-physics community, detailed knowledge about universe could be
incorporated into the world mechanics, from realistic taxonomy and
composition of solar and planetary systems to large-scale structure of
galaxies, galactic clusters and beyond.

General theory of relativity plays an important role in the simulation
of long-range space travel and in presentation of exotic astronomic
phenomena while quantum mechanics and nano-technology present the
basic sciences for future technological development. Additionally,
hypothetical couplings between general relativity and quantum
mechanics could be stipulated and explored in semi-realistic
technological contexts leading to, for example, faster then light
travel and space-time manipulation devices.

Many disciplines of computer science naturally permeate all elements of
the world simulation software and of distributed information services
required for world operation and maintenance. Staying within the
aspects of the world mechanics and presentation, the most important
fields are computer graphics, computer vision, robotics, artificial
intelligence and control of autonomous agents. All of them can be
employed on both sides of the world, by the users to steer their
development and by the world maintainers to control non-interactive
world entities.

Resource acquisition, resource processing and industrial production
can be simulated by various aspects of geology, metallurgy, chemistry,
material science, engineering and process control. Futuro-realistic
simulation of energy generation and consumption provides an
interesting platform for exploration of alternative energy sources,
energy storage technologies and conversion devices. Construction and
operation of infrastructure, habitats and transport vehicles depends
on all of these disciplines, thus allowing the introduction of
complex world constructs, mechanisms and constraints.

Distribution of resources on the level of planetary and solar systems
determines the location of initial settlements and leads, as
technology progresses, to the need for safe and energy-efficient
transportation of raw-materials, products and world entities among
different communities. Trade and other forms of economic cooperation
naturally arise in such circumstances, adding a new level of activity
into the world mechanics. Of course, technology, knowledge and
expertise can likewise become the objects of trade.

Management of human population initially starts with a basic
population dynamics. As human specialists play an important role in
research, engineering and process control, the main emphasis is put
on education and general well being of human communities. Cultural
aspects can be introduced from two sides, from the direction of the
individual and from the direction of the society. In the first case
art and philosophy influence the possibilities of individuals for
attainment of knowledge, expertise and wisdom. In the second case,
human liberties, moral values, ideologies and religions determine the
level of social coercion and define the possible scope of social
manifestations as well as of reactions of an individual to the imposed
conditions. In an extreme case, these models could be used to explore
alternative forms of government, social solidarity and economic
relations with the goal of increasing human prosperity, stimulating
the growth of knowledge and maximizing the rate of technological
development.


\subsection{Levels of realism in \greed worlds}

Combination of all the above elements obviously leads to a world that
is overly complex and, as it seems at the moment, quite impossible to
construct and to maintain. While this remains the final goal, it seems
a sound tactics to start with several disparate worlds that are
constructed from a subset of carefully selected elements, chosen by
considering several aspects of the world simulation.
\begin{itemize}\tightlist
\item \emph{Scope and detail of scientific, technological and
    sociological elements} determines the active world mechanisms and
  sets the main roles available to users.
\item \emph{Intended public} directly influences the
  complexity of both the simulated world and world mechanisms.
\item \emph{Spatial extent, duration and tempo} pose the requirements
  for the server infrastructure. They also determine the amount of
  time users must invest into the interaction with the world to
  accomplish sizable advancement.
\end{itemize}
Users' feedback and interest in actual running worlds will help
to determine what types of worlds are of interest for different
audiences and will further influence the development of core world
simulation framework.

\paragraph{Basic worlds} are built around a small number of basic
elements. The world duration ranges from hours to weeks, the
spatial extent is limited accordingly. They have three main uses:
\begin{enumerate}\tightlist
\item worlds for new users that need to gain initial understanding of
  the world mechanisms and concepts or want to explore a new type of
  world;
\item worlds with content and complexity suitable for children (user
  interface can be simplified to the level of simple computer games);
  and
\item testing of new world mechanisms and their balancing against
  existing elements.
\end{enumerate}
As these worlds are relatively short-lived, they can be regularly
tailored to serve the changing needs of the community.

\paragraph{Intermediate worlds} combine more elements on a higher
level of detail and their duration ranges from months to years. These
worlds can focus on a variety of topics from the repertoire described
in previous sections and only experimentation can show which areas
will attract most users. As the chosen emphasis also determines the
level of users' engagement, several different worlds will be needed to
cover the expectations of individual users.

\paragraph{Realistic worlds} with all available complexity in certain
field can be used by specialists for testing of real-world algorithms,
for development of AI agents and for simulations of space exploration
missions. By keeping the user interface consistent with the one of
intermediate worlds, the users can intuitively enter them to gain
insight into current research or use them as a knowledge data-base
system.


\subsection{Conclusion}

Due to the multitude of directions the development might take, it
seems pointless to consider the final structure and content of the
worlds in a more detailed manner. However, it is important that the
actual flow of development remains influenced by world creators and
users alike, allowing all parties to remain satisfied with the
contract.


% ========================================================================

\section{The \greed partnership}

The \greed partnership is formed among two main parties: the research
institutions and the Internet users, with \greed as an organization
playing the role of the middle man. Initially, \greed will take a form
of a virtual organization for popularization of science and technology
with strong emphasis on dissemination of contemporary information
technologies used in advanced research environments, particularly in
the HEP community. With time it can become a non-profit organization
with the eventual income being spent for further development of \greed
worlds, for maintenance of existing infrastructure and for financing
of projects related to spreading of universal scientific education.

This section analyzes benefits and concessions of different parties
that can join the \greed partnership. All these are, at the end,
a question of subjective rationalization: some seem realistic and
sound while others don't have any appeal at all. Therefore, let
thousand flowers bloom, let every ear hear its story.


\subsection{Internet users / General public}

\paragraph{Benefits:}
\begin{itemize}\tightlist

\item Participation in \greed worlds and communities.

\item Access to knowledge data-bases.

\item Understanding of contemporary research activities and technology.

\item Long-term benefits of increased public interest in scientific
  development and universal education. This is, in fact, the ultimate
  motive of the \greed project.

\end{itemize}

\paragraph{Possible concessions:}
\begin{itemize}\tightlist

\item CPU power, disk, bandwidth.

\item Involvement in world creation and world maintenance.

\item Work for individual sub-projects (core development,
  art-work\footnote{E.g., textures, 3D models, animations, sound
    effects, music, or story-telling.}, or community support).

\end{itemize}


\subsection{Institutions for research and higher education}

\paragraph{Benefits:}
\begin{itemize}\tightlist

\item Distributed computing environment and resources for execution of
  simulations and analyses with high demand for computing resources.

\item \greed worlds provide an environment for construction of complex
  knowledge models and databases that can be utilized and presented on
  different levels. They can be used for research, knowledge
  preservation and for educational purposes.

\item Popularization of individual areas of science by exposing
  scientific models and current research via \greed world mechanisms
  and databases.

\end{itemize}

\paragraph{Possible concessions:}
\begin{itemize}\tightlist

\item Computing resources for other projects. These resources can be
  claimed back from other \greed members when the need arises.

\item Servers for running of \greed worlds and storage of world data.

\item Partake in \greed software development; provide developers for
  core \greed environment or for specific extensions of world
  mechanicsms and knowledge databases.

\end{itemize}


\subsection{Commercial corporations}

Until now, commercial corporations have not been mentioned as possible
members of the \greed partnership. As \greed is, in its essence, a
public welfare project, sponsorship from corporations could be
expected at some basic level. However, there is a possibility for a
far more specific interest of corporations: advertising. Not only
could the advertisements be placed on the sites of \greed worlds as
they are placed on more and more community portals, they could also be
actively introduced as world entities, e.g., Airbus / Boeing
space-ships, Hyundai vehicles, or Novartis life extension drugs. To
prevent complete commercialization of \greed, the participation of
commercial entities should be restricted to a subset of worlds that
offer additional benefits to the users in exchange for being targets
of advertising campaigns.

\paragraph{Benefits:}
\begin{itemize}\tightlist

\item Advertising, in both senses mentioned above. First, specific
  product placement through advertising on community sites, and,
  second, boosting brand and trademark recognition through in-world
  products or entities.

\item Image building by presentation of technological vision or display
  of interest for general progression of technology and science.

\end{itemize}

\paragraph{Possible concessions:}
\begin{itemize}\tightlist

\item Financial support for specific projects or \greed maintenance.

\item Provision of computing resources for operation of \greed worlds.

\item Sponsoring user groups that tie themselves to an interest that
  is close to company's spirit.

\item Providing in-world or real-world rewards for users that meet
  certain criteria or deliver best solutions to calls for creation and
  improvement of in world entities or mechanisms.

\end{itemize}

\emph{In-game advertising} of real-world products and brands has been
used in several titles since the mid-90's. Significant income started
to be gained during the last few years, with the ability to deliver
targeted dynamic advertisements into the on-line virtual worlds and
games.


\subsection{Conclusion}

There are many appeals in the \greed project for all three parties
discussed in this section. A careful analysis confronting the expected
manpower investment against public interest it is expected to arouse
should be made in order to determine the best strategy for placement
of specific options for contribution and benefit of individual
parties. When preparing a public version of this document, the focus
must be known quite specifically and document's structure revised
accordingly.


% ========================================================================


\section{Closing remarks}

In its essence, this document is an invitation to join in early
discussions about \greed, its basic premises and its possible
evolutions. A more concrete technological overview and a proposal
towards the prototype implementation is available on the
web.\footnote{\url{http://www.gled.org/docs/greed-design-docs/greed-tech-workplan.html}}
A community site for wider audience is also being
prepared.\footnote{\url{http://greed.gled.org/}}

It is too early to plan the details and assign commitments. But it is
the right time to consider the possibilities of \greed in the present
reality and to ponder whether the project could fly, regardless what
final form it might take.


\newpage
%%======================================================================
\appendix

\section{Definitions}

\begin{multicols}{2}

\def\ddd#1{\smallskip\noindent \textbf{#1}\quad}

\ddd{\greedhome} Sub-project of \greed that harvests the
computing resources of Internet users.

\ddd{\greed members} Internet users, research institutions, \greed
world developers and commercial corporations that willfully join the
\greed partnership.

\ddd{\greed users} Internet users that contribute their computing
resources and participate in \greed worlds.

\ddd{\greed-world} Sub-project of \greed for creation, maintenance and
operation of virtual worlds available to \greed members.

\ddd{\greed worlds} Virtual worlds simulating scientific and
technological exploration of different universes, ranging from
completely imaginary ones to those that attempt to mimic our reality.

\ddd{world elements} Stars, planets, entities, resources,
infrastructure, technologies, algorithms.

\ddd{world entities} Physical world objects like terrain structures,
buildings, factories, laboratories, vehicles, AI agents and human specialists. Users
can interact with them and influence their future state or actions.

\end{multicols}


% ========================================================================

\newpage

\tableofcontents

% ========================================================================
\end{document}
% ========================================================================


\newpage
%%======================================================================

\section{Ode to Greed}

\begin{verse}
\makebox[16mm][l]{Greed!}   The prime incentive, the dawn of emerging life.\\
\makebox[16mm][l]{Passion!} The mother of change, the queen of ChaOss.\\
\makebox[16mm][l]{Love!}    The source of wisdom, the mirror of truth.
\end{verse}

We follow You the emotions of the heart, You are guiding us where we
have come from, where there is no space nor time, where divine grace
awaits us, its children, its affirmation in the echo of eternity. Out
of mechanical plant-likeness, out of carnal animality, we, the
people of Earth, have been growing forward through the time of our
world, back towards the manifest creation, towards eons born from the
oneness of the unspeakable.

We came to know fire, we came to know the bitterness of the sharpened
metal and we came to know death in the name of God.

We came to know measure and number, we came to know the laws of
heavenly spheres and we came to know the sweet smell of gold.

We came to know the relativity of space and time, we came to know the
uncertainty of quantum worlds and we came to know the two-edged sword
of global information technologies.

Let us cut the Gordian knot of biological evolution!

Let us unshackle the horizons of our existence!

Let us create God the creation has denied us!


\section{Oda Pohlepu}

\begin{verse}
\makebox[18mm][l]{Pohlep!}   Osnovno gibalo, svit porajajoèega se ¾ivljenja.\\
\makebox[18mm][l]{Strast!}   Mati spremembe, kraljica ChaOssa.\\
\makebox[18mm][l]{Ljubezen!} Izvir modrosti, ogledalo resniènosti.
\end{verse}

Sledimo Vam, èustva srca, v\'odite nas tja od koder smo pri¹li, kjer
ni prostora in kjer ni èasa, kjer nebe¹ka milost nespremenjena èaka na
nas, na svoje otroke, na svojo potrditev v odmevu veènosti.  Iz
mehaniène rastlinskosti, iz mesene ¾ivalskosti, smo rastli ljudje
Zemlje, naprej skozi èas na¹ega sveta, nazaj proti razodeti
stvarnosti, proti eonom rojenim iz enosti neizrekljivega.

Spoznali smo ogenj, spoznali smo bridkost nabru¹ene kovine in spoznali
smo smrt v imenu Boga.

Spoznali smo mero in ¹tevilo, spoznali smo zakone nebe¹kih sfer in
spoznali smo sladkasti vonj zlata.

Spoznali smo relativnost prostora in èasa, spoznali smo nedoloèenost
kvantnih svetov in spoznali smo dvorezni meè globalnih informacijskih
tehnologij.

Presekajmo tedaj gordijski vozel biolo¹ke evolucije!

Razklenimo horizonte lastnega obstoja!

Ustvarimo Boga, ki nam ga stvarstvo ni volilo nakloniti!


%%======================================================================
\newpage

\section{Random fragments}

\subsection{about game}

imagination and reality, fiction and materiality, combining in
thousands of ways, allowing thousand flowers to blossom (Mao, what is
the correct wording?), providing inspiration and satisfaction alike
through unbridled sprees of creativity, playfulness and dedication.

linguistics, language theory, true AI in world that is neither ours
nor theirs (of the mathematical beings), but is a shared world where
both genuses can see the same, act the same, speak the same and, maybe
also, think the same. Or maybe not ...  swarm intelligence


\subsection{sort of conclusion}

The \emph{Large Hadron Collider} and its experiments can, without
exaggeration and pomposity, be called the latest wonder of the world,
comparable in intention and scope to the Great pyramids of Giza: to
glorify the truth and to point towards the realm of eternity. Never
before have the scientific and philosophical quests for truth and
purity been further apart from the common world, the world of
greed and lust, the world stigmatized by the global rush for
economic prosperity, for elementary domination that goes beyond caste,
race and nationality, abolishing(?) unsuccessful human beings into hell
of lowest motivations and ambitions, deep into the abyss of
intellectual, emotional and even physical starvation.

But really really \dots\ people should realize the essential role of science
and education \dots\ of far reaching technological enterprises \dots\
of benevolent cooperation. People should dig that abandoning greed is
the first step towards deliverance, towards the Kingdom of Heaven.
